Effective multi-body SU($N$)-symmetric interactions of ultracold fermionic atoms on a 3-D lattice

### 2019-01-29 multi_body_interactions/papers/theory/paper.tex
modified wording in text to minimize the number of revisions that are not made in response to referee comments

As a last comment, we note that our chosen method for constructing an
effective Hamiltonian is distinct from adiabatic elimination methods
which are commonly used in the atomic physics and quantum optics
communities to develop effective theories for e.g.~the low-lying
levels of a Lambda system \cite{brion2007adiabatic,
  james2007effective, reiter2012effective, sanz2016adiabatic}.  Unlike
the perturbatively expressed, but otherwise exact Schieffer-Wolff
transformation, adiabatic elimination methods use approximations which
rely on the fast dynamics of excited states.  While generally
reasonable, these approximations must be made carefully to avoid
potential problems with self-consistency (see section 3 of
ref.~\cite{brion2007adiabatic}), and yield no obvious or
straightforward means to compute effective corrections beyond second
order in the couplings between low- and high-energy sectors of a
Hilbert space \cite{james2007effective, reiter2012effective,
  sanz2016adiabatic}.  While at least one attempt at systematically
computing higher-order corrections in the framework of adiabatic
elimination has recently been made \cite{sanz2016adiabatic}, the
resulting expressions do not lend themselves as nicely to analytical
treatment, and were in any case found by the authors to be equivalent
to a Schrieffer-Wolff transformation.

@@

The form of the bare two-body interaction Hamiltonian $H_{\t{int}}$ in
\eqref{eq:H_int} motivates a diagrammatic representation of terms in
the effective Hamiltonian $H_{\t{eff}}$ in \eqref{eq:H_int_eff},
similarly to the diagrams used to represent elements of the scattering
matrix in field theory.  Effective $M$-body interaction terms at order
$p$ in $H_{\t{int}}$ can be represented by graphs containing $p$
internal vertices.  Each internal vertex and its associated edges
correspond respectively to a coupling constant and the associated
field operators in $H_{\t{int}}$.  An example 2-vertex diagram
representing an effective 3-body interaction term is provided in
figure \ref{fig:diagram}.  All diagrams are read from left to right to
construct a sequence of operators from right to left; the internal
vertices of a diagram are thus strictly ordered, with the $n$-th
internal vertex from the left corresponding to the $n$-th interaction
Hamiltonian $H_{\t{int}}$ from the right in \eqref{eq:H_int_1_2} or
\eqref{eq:H_int_3}.  Solid (dashed) lines represent field operators
acting on the lowest (arbitrary) motional states.  Spatial overlap
factors at each vertex are determined by the motional states of the
edges which connect to (i.e.~field operators associated with) that
vertex.  While it is possible to construct explicit rules for
determining the energetic suppression factors (i.e.~from $\I$) of the
term represented by a diagram, these factors are most easily
determined by examination of the effective Hamiltonians in
\eqref{eq:H_int_1_2} and \eqref{eq:H_int_3}.

>>>>

As a last comment, we note that our chosen method for constructing an
effective Hamiltonian is distinct from adiabatic elimination methods
which are commonly used in the atomic physics and quantum optics
communities to develop effective theories for e.g.~the low-lying
levels of a Lambda system \cite{brion2007adiabatic,
  james2007effective, reiter2012effective, sanz2016adiabatic}.  Unlike
the perturbative, but exact Schieffer-Wolff transformation, adiabatic
elimination methods use approximations which rely on the fast dynamics
of excited states.  While generally reasonable, these approximations
must be made carefully to avoid potential problems with
self-consistency (see section 3 of ref.~\cite{brion2007adiabatic}),
and yield no obvious or straightforward means to compute effective
corrections beyond second order in the couplings between low- and
high-energy sectors of a Hilbert space \cite{james2007effective,
  reiter2012effective, sanz2016adiabatic}.  While at least one attempt
at systematically computing higher-order corrections in the framework
of adiabatic elimination has recently been made
\cite{sanz2016adiabatic}, the resulting expressions do not lend
themselves as nicely to analytical treatment, and were in any case
found by the authors to be equivalent to a Schrieffer-Wolff
transformation.

@@

The form of the bare two-body interaction Hamiltonian $H_{\t{int}}$ in
\eqref{eq:H_int} motivates a diagrammatic representation of terms in
the effective Hamiltonian $H_{\t{eff}}$ in \eqref{eq:H_int_eff},
similarly to the diagrams used to represent elements of the scattering
matrix in field theory.  Effective $M$-body interaction terms at order
$p$ in $H_{\t{int}}$ can be represented by directed graphs containing
$p$ internal vertices.  Each internal vertex and its associated edges
correspond respectively to a coupling constant and the associated
field operators in $H_{\t{int}}$.  An example 2-vertex diagram
representing an effective 3-body interaction term is provided in
figure \ref{fig:diagram}.  All diagrams are read from left to right to
construct a sequence of operators from right to left; the internal
vertices of a diagram are thus strictly ordered, with the $n$-th
internal vertex from the left corresponding to the $n$-th interaction
Hamiltonian $H_{\t{int}}$ from the right in \eqref{eq:H_int_1_2} or
\eqref{eq:H_int_3}.  Solid (dashed) lines represent field operators
acting on the lowest (arbitrary) motional states.  Spatial overlap
factors at each vertex are determined by the motional states of the
edges which connect to (i.e.~field operators associated with) that
vertex.  While it is possible to construct explicit rules for
determining the energetic suppression factors (i.e.~from $\I$) of the
term represented by a diagram, these factors are most easily
determined by examination of the effective Hamiltonians in
\eqref{eq:H_int_1_2} and \eqref{eq:H_int_3}.
